---
title: "Common Workflow Language R"
output:
  BiocStyle::html_document:
    toc: true
    number_sections: true
    highlight: haddock
    css: style.css
    includes:
      in_header: logo.md
---


<!--
%\VignetteIndexEntry{Common Workflow Language R}
%\VignettePackage{cwl}
%\VignetteEngine{knitr::rmarkdown}
-->

```{r style, echo = FALSE, results = 'asis'}
BiocStyle::markdown(css.files = "custom.css")
```

<!-- override white background for classless code chunks -->
<style type="text/css" scoped>
  pre:not([class]) { background-color: #F8F8F8; }
  pre code { background-color: transparent; }
</style>

```{r code, echo = FALSE}
code <- function(...) {
    cat(paste(..., sep = "\n"))
}

code2 <- function(...) {
    cat(paste("```markdown", ..., "\n", "```", sep = "\n"))
}
```

# Introduction
[Common workflow languange](https://github.com/common-workflow-language/common-workflow-language) is a community wise effor to create specifications that enable reproducible, portable data analysis flow. This R package is based on [draf2](http://common-workflow-language.github.io/draft-2/), full details are described in the website. 

This package provides an R interface to create JSON/YAML file to be used in other implementation, for example, [rabix](https://www.rabix.org/) provides very handy interface to desctibe your tool and flow with graphical editor to make development so much easier, it also has local executor as well. 

# Class
This package defines a set of S4/Reference class that represent data model and tool/flow model specified in draft2, in order to enable object-oriented programming in R. Identical class name and fields names in the CWL specification are used, and manuals are directly copied to each class to make reference checking more easier.

The top class is *CWL* class, it comes with soeme utility function that convert the object into list, JSON, YAML. 


## Constructor
To construct the object, you can simply check out the class name (same as the generator name) in the terminal or check its manual to see what fields it need.

```{r}
library(cwl)
InputParameter
```
Then construct it with the constructor directly

```{r}
ip <- InputParameter(id = "bam", type = "File", streamable = TRUE, 
               label = "Bam fles", default = "./test.bam")
```

## Methods
### show method
The first thing you need to pay attention to is that, *show* method is overrided, by default it output YAML which is easier to read. It automatically parse it recurdively. 

```{r}
ip #same as ip$show()
```
You can also show it as JSON

```{r}
ip$show("JSON") #same as ip$show()
```

The default can always be called
```{r}
showDefault(ip)
```
### Convert to list
This packaeg defines a S4 method called `asList`, you can use it for any object, and it's different from `as.list`. `asList` automatically parse any object recursively to turn S4, R5 reference object or simple object into a list, it only parses the fields or slots in the object. 

```{r}
A <- setClass("A", slots = list(x = "character", y = "numeric"))
a <- A(x = "a", y = 3)
## cannot run as.list(a)
asList(a)
asList(ip) # same as ip$toList()
```

And *CWL* class extends to almost all classes defined in this pacakge, so they all comes with the method `obj$toList()`

### Convert to YAML

To convert an object into YAML is also as easy as list, just use S4 method `asYAML` for any object or `obj$toYAML()` for *CWL* extended object. 

```{r}
ip$toYAML()
```

### Convert to JSON

To convert an object into JSON is also as easy as list, just use S4 method `asYAML` for any object or `obj$toJSON()` for *CWL* extended object. 

```{r}
ip$toJSON()
ip$toJSON(pretty = TRUE)
```

# A CommandLineTool example
Here I let's work through an example for a command line 

```
samtools sort -o output.bam -l 1
```

And let's expose the `-l -o` interface into the CWL object. 

Define input parameteres list

```{r}
## long form
ipl <- InputParameterList(
    InputParameter(id = "bam",
                   type = "File",
                   label = "Bam file",
                   description = "Input bam file",
                   inputBinding = CommandLineBinding(
                       position = 1L,
                       separate = TRUE
                   )),
    InputParameter(id = "level",
                   type = "Integer",
                   label = "Compression Level",
                   description = "Set compression level, from 0 (uncompressed) to 9 (best)",
                   inputBinding = CommandLineBinding(
                       position = 2L
                   )),
    InputParameter(id = "prefix",
                   type = "String",
                   label = "Prefix",
                   description = "Write temporary files to PREFIX.nnnn.bam",
                   inputBinding = CommandLineBinding(
                       position = 3L
                   ))
)

```

Define output parameter list

```{r}
opl <- OutputParameterList(
    CommandOutputParameter(
        id = "sorted",
        type = "File",
        outputBinding = CommandOutputBinding(
            glob = "*.bam"
        )
    )
)
```

Wrap them into CommandLineTool object

```{r}
cwl <- CommandLineTool(id = "samtools-sort",
                label = "Samtools sort subcommand",
                description = "Samtools sort: sort bam into sorted bam : )",
                baseCommand = "samtools sort",
                arguments = "out.bam",
                inputs = ipl,
                outputs = opl)
```

Let's see it in JSON

```{r}
cwl$toJSON(pretty = TRUE)
```

Rabix package extends this package, with it's own specific specs, and has shorthands and an easier interface to describe tools. And the web interface allow you to create/describe tools/workflows much more easier. You can also simply import an JSON file to the interface. Please check the *rabix* package for more details. 

